---
title: "Bladder.Affymetrix.2018"
author: "Robert W. Murdoch"
date: "October 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a full pipeline for importaing affymetrix .CEL data (human microarray chip data)
and running full pairwise differential abundance tests.  A full annotation table is also
referenced in order to link probe names to gene names.  

The .CEL data is read and analyzed by the "limma" package

## initializing limma and dependent packages and importing the annotation table

First, the necessary libraries are activated and the annotation table 
(downloaded from an Affymetrix website) is read in:

```{r, echo=FALSE}
library(limma)
library(statmod)
library(affy)
library(affycoretools)
library(oligo)
library(hta20transcriptcluster.db)
```
```{r}
Annotation <- read.csv("HTA-2_0.na35.2.hg19.transcript.csv/HTA-2_0.na35.2.hg19.transcript.csv", header=TRUE, skip=19)
```

The annotation table is very large and many columns are merged redundant database references (using "//" as separator):
```{r}
head(Annotation)
```

## Test matrices

Differential abundance can only be performed on pairs of sample sets.  

```{r pressure}
targets <- read.csv("targets.txt", as.is=T)
targets.tGF.cGF <- read.csv("targets.GFonly.txt", as.is=T)
targets.tGFR.cGFR <- read.csv("targets.GFRonly.txt", as.is=T)
targets.cGFR.cGF <- read.csv("targets.noDrug.Strain.txt", as.is=T)
targets.tGFR.tGF <- read.csv("targets.drugonly.txt", as.is=T)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
